---
title: "Hybrid OA Dashboard"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    css: !expr subugoetheme::pkgdown_template()$assets$css
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(viridis)
library(plotly)
library(shiny)
library(scales)
library(DT)
library(writexl)
library(echarts4r)
library(hoad)
library(forcats)
library(dplyr)
library(stringr)
```

```{r global}
# access to data
hybrid_df <- readr::read_csv(file = path_extdat("hybrid_publications.csv")) %>%
  mutate(license = fct_infreq(license)) %>%
  mutate(publisher = fct_infreq(publisher)) %>%
  mutate(year = factor(issued, levels = c("2013", "2014", "2015", "2016", "2017", "2018", "2019", "2020"))) %>%
  mutate(hybrid_type = ifelse(hybrid_type == "SCOAP", "SCOAP<sup>3</sup>", hybrid_type)) %>%
  mutate(hybrid_type = factor(hybrid_type, levels = c("Open APC (TA)", "Open APC (Hybrid)", "SCOAP<sup>3</sup>"))) %>%
  mutate(domain = paste(domain, suffix, sep = ".")) %>%
  arrange(desc(yearly_publisher_volume))
unpaywall_df <- readr::read_csv(file = path_extdat("unpaywall_df.csv")) %>%
  mutate(year = factor(year, levels = c("2013", "2014", "2015", "2016", "2017", "2018", "2019", "2020")))
```

Overview
=================================================================

Inputs {.sidebar}
-----------------------------------------------------------------------


```{r}
# from https://stackoverflow.com/questions/42148177/how-can-i-build-multiple-inputs-into-my-shiny-app-with-both-updating-and-multipl
# TODO factor this out
selectizeInput(
  inputId = "publishers",
  label = "Selected Publishers",
  choices = c(All = "", levels(hybrid_df$publisher)),
  multiple = TRUE
)
selectizeInput(
  inputId = "journals",
  label = "Selected Journals",
  choices = c(All = "", sort(unique(hybrid_df$journal_title))),
  multiple = TRUE,
  # by default, selection gets cut by 1000
  options = list(maxOptions = length(unique(hybrid_df$journal_title)))
  )


# publisher filter
jn_filtered <- reactive({
  if (is.null(input$publishers)) {
    return(hybrid_df)
  } else {
    return(hybrid_df[hybrid_df$publisher %in% input$publishers, ])
  }
})

observe(updateSelectizeInput(
  session,
  "journals",
  choices = c(All = "", sort(unique(jn_filtered()$journal_title)))
))

# journal_title filter
jn_f <- reactive({
  if (is.null(input$journals)) {
    return(jn_filtered())
  } else {
    return(jn_filtered()[jn_filtered()$journal_title %in% input$journals, ])
  }
})

# prepare dataset for plotting
hybrid_sub <- reactive({
  if (length(unique(jn_f()$publisher)) > 1) {
    hybrid_sub <- jn_f() %>%
      group_by(year, yearly_all, license) %>%
      count() %>%
      mutate(prop = n / yearly_all)
  } else if (length(unique(jn_f()$journal_title)) > 1 &&
    length(unique(jn_f()$publisher)) == 1) {
    hybrid_sub <- jn_f() %>%
      group_by(year, license, yearly_publisher_volume) %>%
      count() %>%
      mutate(prop = n / yearly_publisher_volume)
  } else {
    hybrid_sub <- jn_f() %>%
      group_by(year, license, yearly_jn_volume) %>%
      count() %>%
      mutate(prop = n / yearly_jn_volume)
  }
})
# using unpaywall indicators
# unpaywall <- reactive({
#   hybrid_df %>%
#     filter(journal_title %in% jn_f()$journal_title) %>%
#     group_by(year, journal_title, publisher, jn_y_unpaywall_others) %>%
#     summarise(n = n_distinct(doi_oa)) %>%
#     gather(n, jn_y_unpaywall_others, key = "source", value = "articles") %>%
#     ungroup() %>%
#     group_by(year, source) %>%
#     summarise(articles = sum(articles, na.rm = TRUE)) %>%
#     mutate(
#       source = ifelse(
#         source == "n",
#         "Crossref immediate license",
#         "Other license information\n(Unpaywall)"
#       )
#     )
# })
```

Notice that only those hybrid open access journals were included where
academic institutions sponsored the open access publication according to the [Open APC initiative](https://github.com/openapc/openapc-de)
and where publishers shared license information about immediate open access with [Crossref](https://github.com/CrossRef/rest-api-doc).

Row
-----------------------------------------------------------------------

### Publishers selected

```{r}
renderValueBox({
  publisher_n <- length(unique(jn_f()$publisher))
  valueBox(publisher_n, icon = "fa-filter")
})
```

### Journals selected

```{r}
renderValueBox({
  journal_n <- length(unique(jn_f()$journal_title))
  valueBox(format(journal_n, big.mark = " ", scientific = FALSE),
    icon = "fa-filter"
  )
})
```

### Hybrid OA articles indexed in Crossref

```{r}
renderValueBox({
  hybrid_n <- nrow(jn_f())
  valueBox(format(hybrid_n, big.mark = " ", scientific = FALSE),
    icon = "fa-creative-commons"
  )
})
```


Column {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Hybrid OA licenses found (relative)

```{r}
renderPlotly({
  p <- ggplot(hybrid_sub(), aes(year, prop, fill = license)) +
    xlab("Year") +
    ylab("Hybrid OA / Articles published") +
    scale_fill_manual(
      "License",
      values = colors_license
    ) +
    scale_x_discrete(drop = FALSE) +
    scale_y_continuous(labels = scales::percent) +
    theme_minimal() +
    geom_col(position = position_stack(reverse = TRUE))
  plotly::ggplotly(p)
})
```


### Hybrid OA licenses found (absolute)

```{r}
renderPlotly({
  p <- ggplot(hybrid_sub(), aes(year, n, fill = license)) +
    xlab("Year") +
    ylab("Hybrid OA Articles") +
    scale_fill_manual(
      "License",
      values = colors_license
    ) +
    scale_x_discrete(drop = FALSE) +
    scale_y_continuous(
      labels = function(x) {
        format(x, big.mark = " ", scientific = FALSE)
      }
    ) +
    theme_minimal() +
    geom_col(position = position_stack(reverse = TRUE))
  # p
  plotly::ggplotly(p)
})
```

### Other types of OA license information detected by Unpaywall

```{r}
renderPlotly({
  unpaywall_df() %>%
    filter(journal_title %in% jn_f()$journal_title) %>%
    oa_by_other() %>%
    oa_by_other_plot() %>%
    plotly::ggplotly()
})
```

Row {data-width=400 data-height=350}
-------------------------------------

### Sources of disclosure of OA sponsorship

```{r}
renderPlotly({
  p <- hybrid_df %>%
    filter(journal_title %in% jn_f()$journal_title) %>%
    ggplot(aes(year, ..count.., fill = hybrid_type)) +
    geom_bar(position = position_stack(reverse = TRUE)) +
    xlab("Year") +
    ylab("Articles") +
    scale_x_discrete(drop = FALSE) +
    scale_y_continuous(
      labels = function(x) {
        format(x, big.mark = " ", scientific = FALSE)
      },
      breaks = scales::pretty_breaks()
    ) +
    scale_fill_manual(
      "Sources",
      values = colors_source_disclosure,
      na.value = "#E9E1D7"
    ) +
    theme_minimal()

  tt <- plotly::ggplotly(p, tooltip = c("y"))
  tt$x$data <- lapply(tt$x$data, function(x) {
    x$text <- paste(x$name, x$y, sep = ": ")
    x
  })
  tt
})
```


### Text-mined author emails (first match per article)

```{r}
renderEcharts4r({
  hybrid_df %>%
    filter(journal_title %in% jn_f()$journal_title) %>%
    count(suffix, domain) %>%
    filter(!is.na(domain)) %>%
    e_charts() %>%
    e_treemap(suffix, domain, n,
      leafDepth = "1",
      name = "Corresponding Email"
    ) %>%
    e_tooltip(trigger = "item")
})
```

Compare
=====================================

Row {data-height=250}
--------------------------------------------

### Hybrid open access uptake by Publisher, shown as the proportion of the total number of hybrid open articles found

```{r}
hybrid_all <-
  hybrid_df %>%
  count(year) %>%
  mutate(year = gsub("201", "1", year)) %>%
  mutate(year = gsub("2020", "20", year))
hybrid_df %>%
  mutate(year = gsub("201", "1", year)) %>%
  mutate(year = gsub("2020", "20", year)) %>%
  mutate(publisher_group = fct_lump(publisher, prop = 0.05)) %>%
  group_by(year, publisher_group) %>%
  summarize(n = n()) %>%
  mutate(prop = n / sum(n)) %>%
  mutate(`Proportion in %` = round(prop * 100, 2)) %>%
  ggplot(aes(year, n)) +
  geom_bar(
    data = hybrid_all,
    aes(fill = "All Hybrid OA Articles"),
    color = "transparent",
    stat = "identity"
  ) +
  geom_bar(aes(fill = "by Publisher", label = `Proportion in %`), color = "transparent", stat = "identity") +
  facet_wrap(~publisher_group, nrow = 1) +
  scale_fill_manual(values = c("#b3b3b3a0", "#153268"), name = "") +
  labs(x = "Year", y = "Hybrid OA Articles") +
  theme_minimal() +
  theme(
    legend.position = "top",
    legend.justification = "right"
  ) +
  scale_y_continuous(labels = scales::number_format(big.mark = " ")) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank()) -> publishers_comp_plot

plotly::ggplotly(publishers_comp_plot, tooltip = c("label", "y"))
```

Row {data-height=500}
--------------------------------------------


### Journal hybrid open access share by publisher

```{r}
fillRow(
  flex = c(1, 3),
  inputPanel(
    selectInput(
      inputId = "Year",
      label = "Select Years",
      choices = c(rev(levels(hybrid_df$year))),
      selected = c("2017", "2018", "2019"),
      multiple = TRUE
    ),
    checkboxInput("rb", "Display journals", FALSE)
  ),
  plotlyOutput("pubplot", height = "100%")
)

output$pubplot <- renderPlotly({
  if (input$Year == "All") {
    tt <- hybrid_df %>%
      filter(!is.na(license)) %>%
      mutate(publisher = forcats::as_factor(publisher)) %>%
      group_by(year, publisher, journal_title, yearly_jn_volume) %>%
      summarize(oa = n()) %>%
      mutate(prop = oa / yearly_jn_volume)
  } else {
    tt <- hybrid_df %>%
      filter(!is.na(license)) %>%
      mutate(publisher = forcats::as_factor(publisher)) %>%
      filter(year %in% input$Year) %>%
      group_by(year, publisher, journal_title, yearly_jn_volume) %>%
      summarize(oa = n()) %>%
      mutate(prop = oa / yearly_jn_volume)
  }
  tt <- tt %>%
    ungroup() %>%
    mutate(publisher = forcats::fct_other(publisher, keep = levels(publisher)[1:10]))
  p <- ggplot(tt, aes(y = prop, x = as.factor(publisher))) +
    geom_boxplot() +
    coord_flip() +
    theme_minimal() +
    xlab("Publisher") +
    ylab(NULL) +
    scale_x_discrete(
      drop = FALSE,
      limits = rev(levels(tt$publisher)),
      labels = function(x) {
        str_wrap(x, width = 30)
      }
    ) +
    scale_y_continuous(labels = scales::percent)
  if (input$rb == TRUE) {
    p <- p +
      geom_point(aes(colour = prop, label = journal_title)) +
      scale_colour_viridis(option = "C") +
      scale_fill_viridis(option = "C") +
      guides(color = FALSE, fill = FALSE)
  } else {
    p <- p
  }
  plotly::ggplotly(p, tooltip = c("label", "y")) %>%
    layout(margin = list(l = 200))
})
```


### Journal hybrid open access share by publisher by year

```{r}
fillCol(
  height = 600,
  flex = c(NA, 1),
  inputPanel(
    selectInput(
      "Publisher",
      label = "Select Publisher",
      choices = c("All", unique(hybrid_df$publisher)),
      selected = "All"
    )
  ),
  plotlyOutput("yearplot", height = "100%")
)

output$yearplot <- renderPlotly({
  if (input$Publisher == "All") {
    tt <- hybrid_df %>%
      filter(!is.na(license)) %>%
      group_by(year, publisher, journal_title, yearly_jn_volume) %>%
      summarize(oa = n()) %>%
      mutate(prop = oa / yearly_jn_volume)
  } else {
    tt <- hybrid_df %>%
      filter(!is.na(license)) %>%
      filter(publisher == input$Publisher) %>%
      group_by(year, publisher, journal_title, yearly_jn_volume) %>%
      summarize(oa = n()) %>%
      mutate(prop = oa / yearly_jn_volume)
  }

  p <- ggplot(tt, aes(y = prop, x = as.factor(year))) +
    geom_boxplot() +
    coord_flip() +
    theme_minimal() +
    xlab("Year") +
    ylab(NULL) +
    scale_x_discrete(drop = FALSE) +
    scale_y_continuous(labels = scales::percent)
  if (input$rb == TRUE) {
    p <- p +
      geom_point(aes(colour = prop, label = journal_title)) +
      scale_colour_viridis(option = "C") +
      scale_fill_viridis(option = "C") +
      guides(color = FALSE, fill = FALSE)
  } else {
    p <- p
  }
  plotly::ggplotly(p, tooltip = c("label", "y"))
})
```


Row {data-height=500}
---------------------------------------------

### Table view

```{r}
hybrid_data <- reactive({
  if (input$Publisher == "All") {
    hybrid_df
  } else {
    hybrid_df <- hybrid_df %>%
      filter(publisher == input$Publisher)
  }
  hybrid_df %>%
    filter(year %in% input$Year) %>%
    group_by(year, publisher, journal_title, yearly_jn_volume) %>%
    summarize(oa = n()) %>%
    mutate(prop = round((oa / yearly_jn_volume) * 100, 2)) %>%
    arrange(desc(year)) %>%
    select(
      Year = year,
      Publisher = publisher,
      Journal = journal_title,
      `Article Volume` = yearly_jn_volume,
      `OA Articles` = oa,
      `OA Share (in %)` = prop
    )
})

renderDataTable(
  expr = {
    hybrid_data()
  },
  rownames = FALSE,
  filter = "bottom",
  options = list(
    pageLength = 4,
    dom = "ftp",
    columnDefs = list(list(
      className = "dt-head-left", targets = "_all"
    ))
  )
)
```


Institutional View
================================================================

```{r}
my_data <- reactive({
  if (!is.null(input$plot_clicked_data$name)) {
    hybrid_df %>%
      filter(suffix == input$plot_clicked_data$name | domain == input$plot_clicked_data$name) %>%
      mutate(plot_title = ifelse(!input$plot_clicked_data$name %in% suffix,
        domain,
        suffix
      ))
  } else {
    hybrid_df
  }
})
```

Column
----------------------------------------------------------------------

### Text-mined author emails (first match per article) - Double click for selection!

```{r}
output$plot <- renderEcharts4r({
  hybrid_df %>%
    count(suffix, domain) %>%
    filter(!is.na(domain)) %>%
    e_charts() %>%
    e_treemap(suffix, domain, n,
      leafDepth = "1",
      name = "Corresponding Email"
    ) %>%
    e_tooltip(trigger = "item")
})

output$oa_license <- renderPlotly({
  ggplot(my_data(), aes(factor(year), fill = license)) +
    geom_bar(position = position_stack(reverse = TRUE)) +
    ggtitle(unique(my_data()$plot_title)) +
    scale_fill_manual(
      "License",
      values = colors_license
    ) +
    scale_x_discrete("", drop = FALSE) +
    scale_y_continuous("Articles",
      labels = function(x) {
        format(x, big.mark = " ", scientific = FALSE)
      }
    ) +
    theme_minimal() -> p
  tt <- plotly::ggplotly(p, tooltip = c("y"))
  tt$x$data <- lapply(tt$x$data, function(x) {
    x$text <- paste(x$name, x$y, sep = ": ")
    x
  })
  tt
})

output$oa_source <- renderPlotly({
  p <-
    ggplot(my_data(), aes(factor(year), ..count.., fill = hybrid_type)) +
    geom_bar(position = position_stack(reverse = TRUE)) +
    xlab("Year") +
    ylab("Articles") +
    scale_y_continuous(
      labels = function(x) {
        format(x, big.mark = " ", scientific = FALSE)
      }
    ) +
    scale_fill_manual(
      "Sources",
      values = colors_source_disclosure,
      na.value = "#b3b3b3a0"
    ) +
    theme_minimal()
  tt <- plotly::ggplotly(p, tooltip = c("y"))
  tt$x$data <- lapply(tt$x$data, function(x) {
    x$text <- paste(x$name, x$y, sep = ": ")
    x
  })
  tt
})

fillRow(
  flex = c(1, 2),
  fillCol(echarts4rOutput("plot")),
  fillCol(plotlyOutput("oa_license"), plotlyOutput("oa_source"))
)
```



Row
-------------------------------------------------------------------


### Top Publishers

```{r}
renderDataTable(
  expr = {
    my_data() %>%
      count(publisher) %>%
      arrange(desc(n)) %>%
      mutate(prop = round(n / sum(n) * 100, 2))
  },
  options = list(
    pageLength = 8,
    dom = "ftp"
  )
)
```
