### Open APC Initiative

#### [`oapc_hybrid.csv`](oapc_hybrid.csv)

This dataset was obtained from the [Open APC Initiative](https://github.com/openapc/openapc-de) and was used to determine hybrid open access journals. It also includes data about transformative aggrements, which has no pricing information, as well as country information.

Data schema: <https://github.com/OpenAPC/openapc-de/wiki/schema>

Tibble view in R:

```{r}
library(readr)
readr::read_csv("oapc_hybrid.csv")
```

### Crossref

#### [`jn_facets_df.json`](jn_facets_df.json)

JSON-ND file. Includes the following metadata from Crossref:

|Variable                     |Description                                                        |
|:----------------------------|:------------------------------------------------------------------|
|`issn`                       |ISSNs obtained from Open APC dataset used to query Crossref API    |
|`year_published`             |Lists yearly journal volume obtained from Crossref API facets      |
|`license_refs`               |Obtained Licence URIs                                              |
|`journal_title`              |Most frequent journal title                                        |
|`publisher`                  |Most frequent publisher name                                       |

How to load into R?

```{r eval=FALSE}
library(jsonlite)
jsonlite::stream_in(file("jn_facets_df.json"), simplifyDataFrame = FALSE)
```

#### [`hybrid_license_dois.json`](hybrid_license_dois.json)

JSON-ND file documenting articles found for every journal and open license URI.

Variable                    |Description
|:--------------------------|:--------------------------------------------------------------|
|`dois`                     |List fo DOIs                                                   |
|`issn`                     |ISSN obtained from Open APC dataset used to query Crossref API |
|`license`                  |Open License URI (non-normalised)                              |

```{r eval=FALSE}
library(jsonlite)
jsonlite::stream_in(file("hybrid_license_dois.json"), simplifyDataFrame = FALSE)
```

#### `hybrid_license_md.json`

`data/hybrid_license_md.json`contains full metadata of all open access articles found as it was parsed by the `rcrossref::cr_works()` function. Unfortunately, this data file is simply too large to be shared via GitHub. Please contact me, if you want access to it.

Tibble view in R:

```{r}
library(jsonlite)
library(dplyr)
jsonlite::stream_in(file("../data/hybrid_license_md.json"), verbose = FALSE) %>%
  dplyr::as_data_frame()
```

## Count data

#### [`indicator.csv`](indicator.csv)

[`indicator.csv`](indicator.csv) contains count data about the examined yearly journal volume. 

Coding scheme:

|Variable                    |Description
|:--------------------------|:------------------------------------------------------------------|
|`journal_title`            |Most frequent journal title used by Crossref                       |
|`publisher`                |Most frequent publisher name used by Crossref                      |
|`year`                     |Earliest publishing year                                           |
|`yearly_jn_volume`         |Yearly article volume per journal                                  |
|`license`                  |Open License URI (semi-normalised)                                 |
|`license_ref_n`            |Yearly article volume under the license `license`                  |
|`yearly_publisher_volume`  |Yearly article volume of all journals in the dataset per publisher |   


Tibble view in R:

```{r}
library(readr)
readr::read_csv("indicator.csv")
```

## Flipped journals

To detect fully open acces journals, the [Directory of Open Access Journals](https://doaj.org/) was checked. [`flipped_jns_doaj.csv`](flipped_jns_doaj.csv) contains the so detected articles published in fully open access journals.

```{r}
library(readr)
readr::read_csv("flipped_jns_doaj.csv")
```

Furthermore, [`flipped_jns.csv`](flipped_jns.csv) contains journals that are probably flipped, indicated by a proportion of open access article volume larger than 0.95 in at least two years.

```{r}
library(readr)
readr::read_csv("flipped_jns.csv")
```


## Data re-use and licenses

Open APC Data are made available under the Open Database License: http://opendatacommons.org/licenses/odbl/1.0/. Any rights in individual contents of the database are licensed under the Database Contents License: http://opendatacommons.org/licenses/dbcl/1.0/.